Adaptive Video Streaming with PID-Based Quality Control and Optimized Bitstream Extraction

### 2015-04-05 tmm.tex
add the psnr fluctuation figure

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper proposes an adaptive video streaming system that adjusts the video quality according to available bandwidth and gives the user best quality of service under various network conditions. Two key problems within the system are analyzed and solved with novel algorithms. First, we utilize the Proportional-Integral-Derivative (PID) method in video quality control to determine a suitable bitrate for transmission, so that the system can deliver highest video quality with least quality fluctuation. Second, to select an optimal sub-stream for transmission under given bitrate constraint, we use a linear error model to estimate the distortion caused by discarding any combination of data packets, and a greedy algorithm scheme to assign each data packet a priority value for Rate-Distortion optimized extraction. Experimental results show that the PID-based quality control algorithm brings 8.6\% higher video quality as well as 24.8\% lower quality fluctuation than the baseline algorithm, and the bitstream extraction method achieves a PSNR gain of up to 0.5dB compared to existing reference extractors, with no computational complexity increment. The system is currently running at the online video website www.7dlive.com and proves to be practically effective.
\end{abstract}

@@

\begin{figure}[t]
\centering
\includegraphics[width = 0.9\linewidth]{intervals.jpg}
\caption{Two timelines in Darwin Streaming Sever (DSS) \label{fig:intervals}}
\end{figure}

@@

\begin{figure*}[t]
\subfloat[Average quality level of different control algorithms]{
\includegraphics[width=0.5\textwidth]{quality-all.jpg}
\label{fig:quality-all}}
\subfloat[Variance of quality level of different control algorithms]{
\includegraphics[width=0.5\textwidth]{variance-all.jpg}
\label{fig:variance-all}}
\caption{Comparison of the long/short time prediction and the PDF algorithm in DSS}
\label{fig:performance-all}
\end{figure*}

\begin{figure*}[t]
\centering
\subfloat[Variance of quality level of PID and PDF control algorithms]{
\includegraphics[width=0.5\textwidth]{quality-two.jpg}
\label{fig:quality-two}}
\subfloat[Average quality level of PID and PDF control algorithms]{
\includegraphics[width=0.5\textwidth]{variance-two.jpg}
\label{fig:variance-two}}
\caption{Comparison of the proposed PID algorithm and the PDF algorithm in DDS}
\label{fig:performance-two}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width = 0.9\linewidth]{fluctuation.jpg}
\caption{Video quality fluctuation with time \label{fig:fluctuation}}
\end{figure}

@@

\begin{comment}[8 mini-figures]
\begin{figure*}[t]
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width = \linewidth]{City.jpg}
\caption{City \label{fig:City}}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width = \linewidth]{Foreman.jpg}
\caption{Foreman \label{fig:Foreman}}
\end{minipage}
\end{figure*}

\begin{figure*}[t]
\begin{minipage}{0.5\linewidth}
\centering
\includegraphics[width = \linewidth]{Harbour.jpg}
\caption{Harbour \label{fig:Harbour}}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\centering 
\includegraphics[width = \linewidth]{Soccer.jpg} 
\caption{Soccer \label{fig:Soccer}}
\end{minipage} 
\end{figure*}

\end{comment}

@@

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In order to give the user best quality of service under bandwidth fluctuation, video streaming systems should be able to adjust the video quality according to available bandwidth. In this paper we propose novel models and algorithms to enable such an adaptive video streaming system which adopts the Scalable Video Coding (SVC) technology. First, we utilize the Proportional-Integral-Derivative (PID) method in quality control to determine a suitable transmission bitrate according to actual bandwidth condition. By considering bandwidth information of the past, the current and even the future, the PID-based quality control algorithm is able to deliver highest video quality with least quality fluctuation. Second, we present an efficient bitstream extraction scheme to select an optimal sub-stream from the whole SVC video bitstream under given bitrate constraint. A linear error model is used to accurately estimate the distortion caused by discarding any combination of data packets, and a greedy-like algorithm is designed to assign each data packet a priority value for Rate-Distortion optimized extraction. Experimental results show that comparing with the existing solutions, the PID-based quality control algorithm brings 8.6\% higher video quality with 24.8\% lower quality fluctuation, and the bitstream extraction method achieves a PSNR gain of up to 0.5dB with no computational complexity increment. The proposed algorithms are implemented in online video website www.7dlive.com and prove to be practically effective.
\end{abstract}

@@

\begin{figure}[t]
\centering
\includegraphics[width = 0.9\linewidth]{Intervals.jpg}
\caption{Two timelines in Darwin Streaming Sever (DSS) \label{fig:intervals}}
\end{figure}

@@

\begin{figure*}[t]
\subfloat[Average quality level of different control algorithms]{
\includegraphics[width=0.5\textwidth]{Quality-all.jpg}
\label{fig:quality-all}}
\subfloat[Variance of quality level of different control algorithms]{
\includegraphics[width=0.5\textwidth]{Variance-all.jpg}
\label{fig:variance-all}}
\caption{Comparison of the long/short time prediction and the PDF algorithm in DSS}
\label{fig:performance-all}
\end{figure*}

\begin{figure*}[t]
\centering
\subfloat[Variance of quality level of PID and PDF control algorithms]{
\includegraphics[width=0.5\textwidth]{Quality-two.jpg}
\label{fig:quality-two}}
\subfloat[Average quality level of PID and PDF control algorithms]{
\includegraphics[width=0.5\textwidth]{Variance-two.jpg}
\label{fig:variance-two}}
\caption{Comparison of the proposed PID algorithm and the PDF algorithm in DDS}
\label{fig:performance-two}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width = 0.9\linewidth]{Fluctuation.jpg}
\caption{Video quality fluctuation with time \label{fig:fluctuation}}
\end{figure}

@@

@@

\subsection{Experiments for combined system}

After combining the two algorithm together, we compared the proposed system's performance with the original DSS with JSVM Basic bitstream extraction. The result can be seen in 
Fig. \ref{fig:fluctuation-psnr}.

\begin{figure}[t]
\centering
\includegraphics[width = 0.9\linewidth]{Fluctuation-PSNR.jpg}
\caption{PSNR fluctuation with time \label{fig:fluctuation-psnr}}
\end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width = 0.9\linewidth]{Fluctuation-PSNR-wide.jpg}
\caption{PSNR fluctuation with time \label{fig:fluctuation-psnr-wide}}
\end{figure*}
